\bf Hierarchical Network Structure Promotes Dynamical Robustness

### 2015-07-22 tex/introduction.tex
add relevant citations to introduction

Recent work has demonstrated that as a result of the existence of largely insensitive directions within the parameter space for such models, the approach outlined above often allows for a large variety of models to fit equivalent data \cite{Machta2013,Hines2014,Prabakaran2014,Tonsing2014}. In addition, there is often uncertainty about the very structure of such networks. Since the evolutionary process results in modifications to the underlying model this fact may be used to characterize evolutionarily effective versus neutral spaces. In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems. This is necessary to understand in order to determine whether or not a given dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}. This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems that, in comparison to one another, appear to have components that are randomly interlinked.
% \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}

@@

Investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}. The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the networks represented in \ref{fig:biomodelexamples}. May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014,Majumdar2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.

>>>>

Recent work has demonstrated that as a result of the existence of largely insensitive directions within the parameter space for such models, the approach outlined above often allows for a large variety of models to fit equivalent data \cite{Machta2013,Hines2014,Prabakaran2014,Tonsing2014}. In addition, there is often uncertainty about the very structure of such networks. Since the evolutionary process results in modifications to the underlying model this fact may be used to characterize evolutionarily effective versus neutral spaces. In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems. This is necessary to understand in order to determine whether or not a given dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}. This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems \cite{Bialek2012}. For the case considered here, the ensemble of interest is defined by dynamical systems that, in comparison to one another, appear to have components that are randomly interlinked.
% \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}

@@

Investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}. The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the networks represented in \ref{fig:biomodelexamples}. May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,McMurtrie1975,Radius2014,Majumdar2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.

### 2015-07-22 tex/overview.tex
add relevant citations to introduction

% To determine stability, we linearize the equations of motion (\refsupp{}) about the
% fixed point $\vec{x}^0$:
% \begin{equation}\label{eq:lineardynsys}
% \frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
% \end{equation}
% where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$ Jacobian matrix $A$ has components
% $$
% a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
% $$
\subsection{Stability analysis of biological networks}
To determine stability, we use the Taylor series expansion of the equations of motion \ref{eq:eom} about the fixed point $\vec{x}^0$ where $\vec{y} = \vec{x} - \vec{x}^0$ by
\begin{equation}\label{eq:taylorseries}
\begin{aligned}
\frac{dx_i(t)}{dt} \approx & F_i(\vec{x}^0)
+ \sum_{j=1}^{N} \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0} y_j\\
& + \frac{1}{2}\sum_{j,k=1}^{N} \left. \frac{\partial^2 F_i}{\partial x_j \partial x_k} \right|_{\vec{x} = \vec{x}^0} y_j y_k + \cdots
\end{aligned}
\end{equation}
The zeroth order term vanishes since $F_i(\vec{x}^0)=0$ by definition and thus neglecting terms higher than first order from \ref{eq:taylorseries} results in
% \begin{equation*}
% \frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
% \end{equation*}
\begin{equation}\label{eq:lineardynsys}
\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
\end{equation}
where the $n \times n$-matrix $A$ has components
$$
a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
$$

>>>>

% To determine stability, we linearize the equations of motion (\refsupp{}) about the
% fixed point $\vec{x}^0$:
% \begin{equation}\label{eq:lineardynsys}
% \frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
% \end{equation}
% where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$ Jacobian matrix $A$ has components
% $$
% a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
% $$
\subsection{Stability analysis of biological networks}
To determine stability, we use the Taylor series expansion of the equations of motion, \ref{eq:eom}, about the fixed point $\vec{x}^0$ where $\vec{y} = \vec{x} - \vec{x}^0$ given by
\begin{equation}\label{eq:taylorseries}
\begin{aligned}
\frac{dx_i(t)}{dt} \approx & F_i(\vec{x}^0)
+ \sum_{j=1}^{N} \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0} y_j\\
& + \frac{1}{2}\sum_{j,k=1}^{N} \left. \frac{\partial^2 F_i}{\partial x_j \partial x_k} \right|_{\vec{x} = \vec{x}^0} y_j y_k + \cdots
\end{aligned}.
\end{equation}
The zeroth order term vanishes since $F_i(\vec{x}^0)=0$ by definition and thus neglecting terms higher than first order from \ref{eq:taylorseries} results in
% \begin{equation*}
% \frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
% \end{equation*}
\begin{equation}\label{eq:lineardynsys}
\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
\end{equation}
where the $n \times n$-matrix $A$ has components
$$
a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
$$
